### 2017-05-24 paper/sexSelectionSpacing-ver1.tex
Work on new introduction
- Simple edits of main paper
- New intro in notes.txt

\title{Son Preference, Sex-Selective Abortions,\\ and Birth Spacing
\protect\thanks{%
I am grateful to Andrew Foster and Darryl Holman for discussions about the method.
I owe thanks to three anonymous referees, Shelly Lundberg, Daniel Rees, David Ribar, 
Hendrik Wolff, seminar participants at University of Copenhagen, University of Michigan, 
University of Washington, University of \AA{}rhus, the Fourth 
Annual Conference on Population, Reproductive Health, 
and Economic Development, and the Economic Demography Workshop for helpful suggestions and comments.
I would also like to thank Nalina Varanasi for research assistance.
Support from the University of Washington Royalty Research Fund and the 
Development Research Group of the World Bank is gratefully acknowledged.
The views and findings expressed here are those of the author and
should not be attributed to the World Bank or any of its member countries.
Partial support for this research came from a Eunice Kennedy Shriver National
Institute of Child Health and Human Development research infrastructure grant,
5R24HD042828, to the Center for Studies in Demography and Ecology at the
University of Washington.
% Prior versions of this paper were presented under the title ``The 
% Determinants of sex-selective Abortions.''
}
}

@@

% Demography abstract
\noindent This paper addresses the question:
how does birth spacing interact with the use of sex-selective abortions? 
I introduce a statistical method that incorporates how sex-selective abortions 
affect both the likelihood of a son and spacing between births.
Using India's National Family and Health Surveys,
I show that use of sex selection leads to longer spacing after a daughter is born.
Women with 8 or more years of education, both in urban and rural areas, are the main users 
of sex-selective abortions.
Women with less education do not appear to use sex selection.
Predicted lifetime fertility for high-education women declined more than 10\% 
between 1985--1994, when sex selection was legal, and 1995--2006, when sex selection was illegal.
Abortions per woman increased almost 20\% for urban women and 50\%
for rural women between the two periods, suggesting that making sex selection illegal has 
not reversed its use.
Finally, sex selection appears to be used to ensure one son rather than multiple sons.

@@

Understanding and measuring son preference has been an important research 
area for decades and is as salient as ever.
[brief description of where identified]
Three approaches to measuring son preference have been in widespread use:
birth spacing, parity progression, and sex ratio \citep{Leung1988}.

Before prenatal sex determination became widely available, parents' only 
option if they wanted a son was to try for another pregnancy as soon as 
possible after the birth of a girl.
A preference for sons is therefore often associated with shorter birth spacing 
after the birth of a daughter than after the birth of a son 
\citep{Das1987,Rahman1993,Pong1994,Haughton1996,Arnold1997,Soest2012}.
The resulting shorter spacing is thought to be associated with worse 
health outcomes for the girls 
\citep{arnold98,Whitworth2002,Rutstein2005,Conde-Agudelo2006}.
Parents are also more likely to cease childbearing after the birth of 
a son than after a daughter, resulting in a lower parity progression
with the number of sons \citep{ben-porath76b,Das1987,Arnold1997,clark00}. 
Finally, increase in the ratio of males to females 
can serve as an indication of son preference because 
of differential mortality, for example because of the shorter
birth spacing, or underreporting \citep{sen90,sen92,Merli2000,anderson08}.

>>>>

\title{Son Preference, Birth Spacing, and Sex-Selective Abortions%
\protect\thanks{%
I am grateful to Andrew Foster and Darryl Holman for discussions about the method.
I owe thanks to three anonymous referees, Shelly Lundberg, Daniel Rees, David Ribar, 
Hendrik Wolff, seminar participants at University of Copenhagen, University of Michigan, 
University of Washington, University of Aarhus, the Fourth 
Annual Conference on Population, Reproductive Health, 
and Economic Development, and the Economic Demography Workshop for helpful suggestions and comments.
I would also like to thank Nalina Varanasi for research assistance.
Support from the University of Washington Royalty Research Fund and the 
Development Research Group of the World Bank is gratefully acknowledged.
The views and findings expressed here are those of the author and
should not be attributed to the World Bank or any of its member countries.
Partial support for this research came from a Eunice Kennedy Shriver National
Institute of Child Health and Human Development research infrastructure grant,
5R24HD042828, to the Center for Studies in Demography and Ecology at the
University of Washington.
% Prior versions of this paper were presented under the title ``The 
% Determinants of sex-selective Abortions.''
}
}

@@

% Demography abstract
\noindent This paper examines how birth spacing interact with the use of 
sex-selective abortions. 
I introduce a statistical method that incorporates that sex-selective abortions 
affect both the likelihood of a son and spacing between births.
Using India's National Family and Health Surveys,
I show that use of sex selection leads to \emph{longer} spacing after a daughter 
is born.
Women with 8 or more years of education, both in urban and rural areas, are 
the main users of sex-selective abortions, whereas women with less education 
do not appear to use sex selection.
Predicted lifetime fertility for high-education women declined more than 10\% 
between 1985--1994, when sex selection was legal, and 1995--2006, when sex selection 
was illegal.
Abortions per woman increased almost 20\% for urban women and 50\%
for rural women between the two periods, suggesting that making sex 
selection illegal has not reversed its use.
Finally, sex selection appears to be used to ensure one son rather 
than multiple sons.

@@

The effects of son preference has been an active research area for decades.
Before prenatal sex determination, a preference for sons was often associated 
with shorter birth spacing after the birth of a daughter than after the 
birth of a son, because that was the only way to achieve a son 
\citep{Das1987,Rahman1993,Pong1994,Haughton1996,Arnold1997,Soest2012}.
The resulting shorter spacing is associated with worse health outcomes 
for the girls 
\citep{arnold98,Whitworth2002,Rutstein2005,Conde-Agudelo2006}.
Parents are also more likely to cease childbearing after the birth of 
a son than after a daughter, resulting in a lower parity progression
with the number of sons \citep{ben-porath76b,Das1987,Arnold1997,clark00}. 
Finally, increase in the ratio of males to females 
can serve as an indication of son preference because 
of differential mortality, for example because of the shorter
birth spacing, or underreporting \citep{sen90,sen92,Merli2000,anderson08}.
